Symbolic Abstract Data Type Inference

### 2015-11-02 algorithm.tex
Within the page limits.

%!TEX root = draft.tex
\section{A Symbolic ADT Inference Algorithm}
\label{sec:algorithm}

Algorithm~\ref{alg:inference} solves the symbolic ADT inference problem by
computing a finite representation of an ADT complement $B$ using the $\preceq$
relation of Section~\ref{sec:patterns}. In general, this is achieved by
enumerating the elements of $B$ while maintaining the $\preceq$-minimal
elements, and recognizing a condition under which all the elements of $B$ are
related to the current set of minimals~\cite{conf/lics/AbdullaCJT96,
journals/tcs/FinkelS01}. In our case, we stratify our enumeration of $B$ by the
relative sizes of its histories. Formally, we say the \emph{weight} of a history
is the maximum among operation frequencies and the number of matches. We then
define
\begin{align*}
  & B_i = \set{ h \in B : \weight(h) \le i } \\
  & B_i' = \set{ h \in B : \exists h' \in B_i.\ h' \preceq h \text{ and } \weight(h) \le i+1 }
\end{align*}
respectively as the histories of $B$ with at most $i$ matches and duplicates,
and those derived from $B_i$ with at most $i+1$ matches and duplicates. We say
that an ADT with complement $B$ is \emph{predictable} if $B_i^* = B$ whenever
$B_i' = B_{i+1}$, i.e.,~if all histories of $B$ are represented by $B_i$ whenever
all histories of $B_{i+1}$ are represented by $B_i$. Algorithm~\ref{alg:inference}
then performs a weight-increasing enumeration of $B$, collecting
$\preceq$-minimals from smaller-weight violations before advancing to greater
weights. When no violation is found at a given weight, the algorithm terminates.
This algorithm is guaranteed to terminate since $\preceq$ is a
wqo~\cite{conf/lics/AbdullaCJT96, journals/tcs/FinkelS01}. Furthermore, this
algorithm is sound for arbitrary ADTs, and complete for predictable ADTs.
Many naturally-occurring ADTs are predictable — in fact all of the examples we
know of are predictable, as demonstrated in Section~\ref{sec:nature}.

>>>>

\section{A Symbolic ADT Inference Algorithm}
\label{sec:algorithm}

Algorithm~\ref{alg:inference} solves symbolic ADT inference by
computing a finite representation of an ADT complement $B$ using the $\preceq$
relation of Section~\ref{sec:patterns}. This is generally achieved by
enumerating $B$ while maintaining the $\preceq$-minimal
elements, and recognizing a condition under which all the elements of $B$ are
related to the current set of minimals~\cite{conf/lics/AbdullaCJT96,
journals/tcs/FinkelS01}. In our case, we stratify our enumeration of $B$ by the
relative sizes of its histories. Formally, the \emph{weight} of a history
is the maximum among operation frequencies and the number of matches. We then
define
\begin{align*}
  & B_i = \set{ h \in B : \weight(h) \le i } \\
  & B_i' = \set{ h \in B : \exists h' \in B_i.\ h' \preceq h \text{ and } \weight(h) \le i+1 }
\end{align*}
respectively as the histories of $B$ with at most $i$ matches and duplicates,
and those derived from $B_i$ with at most $i+1$ matches and duplicates. We say
that an ADT with complement $B$ is \emph{predictable} if $B_i^* = B$ whenever
$B_i' = B_{i+1}$, i.e.,~if all histories of $B$ are represented by $B_i$ whenever
all histories of $B_{i+1}$ are represented by $B_i$. Algorithm~\ref{alg:inference}
then performs a weight-increasing enumeration of $B$, collecting
$\preceq$-minimals from smaller-weight violations before advancing to greater
weights. When no violation is found at a given weight, the algorithm terminates.
This algorithm is guaranteed to terminate since $\preceq$ is a
wqo~\cite{conf/lics/AbdullaCJT96, journals/tcs/FinkelS01}. Furthermore, this
algorithm is sound for arbitrary ADTs, and complete for predictable ADTs.
Many naturally-occurring ADTs are predictable — in fact all of the examples we
know of are predictable, as demonstrated in Section~\ref{sec:nature}.

### 2015-11-02 formula.tex
Fixed typos from Reviewer A.
Corrections from Reviewer B.
Within the page limits.

%!TEX root = draft.tex
\section{Symbolic ADT Representations}
\label{sec:formula}

While Section~\ref{sec:patterns} demonstrates that (the complements) of
naturally-occurring ADTs can be represented finitely, in this section we
demonstrate that such finite representations have logical interpretations,
allowing us to derive the formulas representing ADTs. In what follows, we
describe how to obtain a formula which is satisfied by the histories which
embed a given history, via $\preceq$. Then, using the finite set of histories
which represent the complement of a given ADT, we represent the ADT itself as
the conjunction of negations of these embedding formulas. The resulting formula
is satisfied only on the histories which do not embed the generators of a given
ADT’s complement.

Let $h = \tup{O,<,c,f,m,r}$ be a history with operations $O = \set{ o_1, \ldots,
o_n }$. Without loss of generality, we assume the match targets $\set{o_1,
\ldots, o_k}$ of $h$ are indexed consecutively from $1$ to $k$, for some $k \le
n$. For each $o \in O$, we define the macro $\textsc{Embed}_o(x,Y,z)$:
\begin{align*}
  & \big(c(o) \implies {\sf c}(x)\big) \land {\sf f}(x) = f(o) \land {\sf r}(x) \Leftrightarrow r(o) \\
  & \land \big(o \in \dom(m)
    \implies \lnot{\sf um}(x) \land {\sf m}(x) = z\big) \\
  & \land \big(o \not\in \dom(m) \implies {\sf um}(x)\big)
    \land \bigwedge_{y \in Y} x < y
\end{align*}
capturing the correspondence between the operation $o$ and the logical variable
$x$ representing $o$. The variables $Y$ and $z$ represent the operations
ordered after $x$, and the operation which $x$ matches. The macro
$\textsc{Identical}(x,y)$:
\begin{align*}
  & \big({\sf c}(x) \Leftrightarrow {\sf c}(y)\big)
    \land {\sf f}(x) = {\sf f}(y)
    \land \big({\sf r}(x) \Leftrightarrow {\sf r}(y)\big) \\
  & \land \big({\sf um}(x) \Leftrightarrow {\sf um}(y)\big)
    \land {\sf m}(x) = {\sf m}(y)
\end{align*}
captures whether the operations bound to $x$ and $y$ are identical.
To express the constraints among the matches of embedded operations, we define
the macro $\textsc{Match}(Y,z)$:
\begin{align*}
  \forall x.
  {\sf m}(x) = z \implies
  {\sf r}(x) \lor \bigvee_{y \in Y} \textsc{Identical}(x,y)
\end{align*}
which requires any operation which matches $z$ to be either read-only, or
identical to some operation in $Y$, which represents the operations of $h$
which match $z$. Finally, we express the embedding of $h$ with the macro
$\textsc{Embed}_h$:
\begin{align*}
  \exists x_1, \ldots, x_n.
  \bigwedge_{i=1}^{n} \textsc{Embed}_{o_i}(x_i,Y_i,z_i)
  \land \bigwedge_{i=1}^{k} \textsc{Match}(W_i, x_i)
\end{align*}
where $Y_i = \set{ x_j : o_i < o_j }$ are the variables corresponding to
operations ordered after $o_i$, and $z_i$ is the variable corresponding to
$m(o_i)$, if defined, and $W_i = \set{ x_j : m(o_j) = o_i }$ are the variables
corresponding to operations matching $o_i$.

@@

  Consider again the histories of Example~\ref{ex:patterns} which generate the
  complement of the atomic register ADT. The $\textsc{Embed}$ formula for the
  first history,
\begin{verbatim}
  [1:X] read => 1 (RO)  #
\end{verbatim}
  after simplifications, like replacing $\mathrm{true} \implies p$ with $p$, is
  \begin{align*}
    \exists x_1.\ {\sf c}(x_1) \land {\sf f}(x_1) = \mathrm{read} \land {\sf um}(x_1) \land {\sf r}(x_1).
  \end{align*}
  The $\textsc{Embed}$ formula for second history,
\begin{verbatim}
  [1:1] write(1)            #
  [2:2] read => - (RO)    #
\end{verbatim}
  is similarly given by
  \begin{align*}
    & \exists x_1, x_2.\ x_1 < x_2 \\
    & \quad \land {\sf c}(x_1) \land {\sf f}(x_1) = \mathrm{write}
      \land \lnot{\sf um}(x_1) \land {\sf m}(x_1) = x_1 \\
    & \quad \land {\sf c}(x_2) \land {\sf f}(x_2) = \mathrm{read}
      \land \lnot{\sf um}(x_2) \land {\sf m}(x_2) = x_2\land {\sf r}(x_2) \\
    & \quad \land \big(\forall x.\ {\sf m}(x) = x_1 \implies {\sf r}(x) \lor \textsc{Identical}(x,x_1)\big) \\
    & \quad \land \big(\forall x.\ {\sf m}(x) = x_2 \implies {\sf r}(x) \lor \textsc{Identical}(x,x_2)\big).
  \end{align*}
  The formulas for the other histories are similarly obtained.

@@

We obtain a formula representing an ADT by taking the conjunction of the
negative embedding formulas $\set{ \lnot\textsc{Embed}_h : h \in H }$ from any
set $H$ that generates its bounded complement $B$, i.e.,~whose closure under
$\preceq$, denoted by $H^*$ henceforth, is equal to $B$. We define the
\emph{exclusion formula} of a set of histories $H$ as $F(H) = \bigwedge_{h \in
H} \neg \textsc{Embed}_h$.

@@

  For a set $\Sigma$, let $\neg \Sigma$ denote its complement. By
  Lemma~\ref{lem:ADT_closure}, the complement of $A$ is closed under $\preceq$.
  Since $H\subseteq B\subseteq  \neg A \subseteq \neg A$, we have that
  $H^* \subseteq \neg A$. By Lemma~\ref{lem:formula1}, the formula $F(H)$ holds
  exactly for the set of histories $\neg H^*$. By the previous inclusions, we
  have that $A\subseteq \neg H^* \subseteq \neg B$. Therefore, $F(H)$ holds for
  all histories in $A$ and doesn't hold for histories in $B$.
\end{proof}

@@

%For normal and predictable ADTs, such sets are computable, and so our inference
%problem is also computable.
%
%\begin{theorem}
%
%  The symbolic ADT inference problem is computable for normal and predictable ADTs.
%
%\end{theorem}

>>>>

\section{Symbolic ADT Representations}
\label{sec:formula}

While Section~\ref{sec:patterns} demonstrates that (the complements of)
naturally-occurring ADTs can be represented finitely, in this section we
demonstrate that those representations have logical interpretations,
allowing us to derive formulas representing ADTs. In what follows we
describe how to obtain formulas satisfied by the histories which
embed a given history via $\preceq$. Then, using the finite set of histories
which represent the complement of a given ADT, we represent the ADT itself as
the conjunction of negations of these embedding formulas. The resulting formula
is satisfied only on the histories which do not embed the generators of a given
ADT’s complement.

Let $h = \tup{O,<,c,f,m,r}$ be a history with operations $O = \set{ o_1, \ldots,
o_n }$. Without loss of generality, we assume the match targets $\set{o_1,
\ldots, o_k}$ of $h$ are indexed consecutively from $1$ to $k$ for some $k \le
n$. For each $o \in O$, the macro $\textsc{Embed}_o(x,Y,z)$:
\begin{align*}
  & \big(c(o) \implies {\sf c}(x)\big) \land {\sf f}(x) = f(o) \land {\sf r}(x) \Leftrightarrow r(o) \\
  & \land \big(o \in \dom(m)
    \implies \lnot{\sf um}(x) \land {\sf m}(x) = z\big) \\
  & \land \big(o \not\in \dom(m) \implies {\sf um}(x)\big)
    \land \bigwedge_{y \in Y} x < y
\end{align*}
captures the correspondence between the operation $o$ and the logical variable
$x$ representing $o$. The variables $Y$ and $z$ represent the operations
ordered after $x$, and the operation which $x$ matches. The macro
$\textsc{Identical}(x,y)$:
\begin{align*}
  & \big({\sf c}(x) \Leftrightarrow {\sf c}(y)\big)
    \land {\sf f}(x) = {\sf f}(y)
    \land \big({\sf r}(x) \Leftrightarrow {\sf r}(y)\big) \\
  & \land \big({\sf um}(x) \Leftrightarrow {\sf um}(y)\big)
    \land {\sf m}(x) = {\sf m}(y)
\end{align*}
captures whether the operations bound to $x$ and $y$ are identical.
To express the constraints among the matches of embedded operations, we define
the macro $\textsc{Match}(Y,z)$:
\begin{align*}
  \forall x.
  \lnot{\sf um}(x) \land
  {\sf m}(x) = z \implies
  {\sf r}(x) \lor \bigvee_{y \in Y} \textsc{Identical}(x,y)
\end{align*}
which requires any operation which matches $z$ to be either read-only, or
identical to some operation in $Y$, which represents the operations of $h$
which match $z$. Finally, we express the embedding of $h$ with the macro
$\textsc{Embed}_h$:
\begin{align*}
  \exists x_1, \ldots, x_n.
  \bigwedge_{i=1}^{n} \textsc{Embed}_{o_i}(x_i,Y_i,z_i)
  \land \bigwedge_{i=1}^{k} \textsc{Match}(W_i, x_i)
\end{align*}
where $Y_i = \set{ x_j : o_i < o_j }$ are the variables corresponding to
operations ordered after $o_i$, and $z_i$ is the variable corresponding to
$m(o_i)$, if defined, and $W_i = \set{ x_j : m(o_j) = o_i }$ are the variables
corresponding to operations matching $o_i$.

@@

  Consider again the histories of Example~\ref{ex:patterns} which generate the
  complement of the atomic register ADT. The $\textsc{Embed}$ formula for the
  first history,
\begin{verbatim}
  [1:X] read => 1 (RO)  #
\end{verbatim}
  after simplifications, like replacing $\mathrm{true} \implies p$ with $p$, is
  \begin{align*}
    \exists x_1.\ {\sf c}(x_1) \land {\sf f}(x_1) = \mathrm{read} \land {\sf um}(x_1) \land {\sf r}(x_1).
  \end{align*}
  The $\textsc{Embed}$ formula for the second history,
\begin{verbatim}
  [1:1] write(1)            #
  [2:2] read => - (RO)    #
\end{verbatim}
  is similarly given by
  \begin{align*}
    & \exists x_1, x_2.\ x_1 < x_2 \\
    & \quad \land {\sf c}(x_1) \land {\sf f}(x_1) = \mathrm{write}
      \land \lnot{\sf um}(x_1) \land {\sf m}(x_1) = x_1 \\
    & \quad \land {\sf c}(x_2) \land {\sf f}(x_2) = \mathrm{read}
      \land \lnot{\sf um}(x_2) \land {\sf m}(x_2) = x_2\land {\sf r}(x_2) \\
    & \quad \land \big(\forall x.\ {\sf m}(x) = x_1 \implies {\sf r}(x) \lor \textsc{Identical}(x,x_1)\big) \\
    & \quad \land \big(\forall x.\ {\sf m}(x) = x_2 \implies {\sf r}(x) \lor \textsc{Identical}(x,x_2)\big).
  \end{align*}
  The formulas for the other histories are similarly obtained.

@@

We obtain a formula representing an ADT by taking the conjunction of the
negative embedding formulas $\set{ \lnot\textsc{Embed}_h : h \in H }$ from any
set $H$ that generates its bounded complement $B$, i.e.,~whose closure under
$\preceq$, denoted by $H^*$ henceforth, is equal to $B$. The
\emph{exclusion formula} of a set of histories $H$ is $F(H) = \bigwedge_{h \in
H} \neg \textsc{Embed}_h$.

@@

  Let $\neg \Sigma$ denote the complement of a set $\Sigma$. By
  Lemma~\ref{lem:ADT_closure}, the complement of $A$ is closed under $\preceq$.
  Since $H^* \subseteq B\subseteq \neg A$, we know
  $H^* \subseteq \neg A$. By Lemma~\ref{lem:formula1}, the formula $F(H)$ holds
  exactly for the histories $\neg H^*$. By the previous inclusions, we
  have that $A\subseteq \neg H^* \subseteq \neg B$. Therefore, $F(H)$ holds for
  $A$ and not for $B$.
\end{proof}

@@

### 2015-11-02 matching.tex
Corrections from Reviewer B.
Within the page limits.

%!TEX root = draft.tex
\section{Matching Scheme Inference}
\label{sec:matching}

@@

The required manual specification includes identifying a class of executions to
which an implementation should be subjected, and a set of predicates that are
required in the logic of matching schemes. Formally, a \emph{language} $L =
\tup{E,P,Q}$ is a set $E$ of executions, along with finite sequences $P = P_1
P_2 \ldots$ and $Q = Q_1 Q_2 \ldots$ of binary and ternary predicates
$P_i(e,o_1)$ and $Q_i(e,o_1,o_2)$ ranging over the executions of $E$ and their
operations. When $P$ is a $k$-length sequence of $n$-ary predicates, we write
$P(x_1, \ldots, x_n)$ to denote the valuation sequence
\begin{align*}
  P_1(x_1, \ldots, x_n) \ldots P_k(x_1, \ldots, x_n).
\end{align*}
Given a language $L = \tup{E,P,Q}$, we say a matching scheme $M$ is
\emph{simple} when there exists an $n$-ary Boolean function $G$, for $n =
2\cdot|P|+|Q|$, such that for each execution $e \in E$
\begin{itemize}

@@

  We say an execution of read and write methods \emph{writes unique values} if
  the argument value to each write operation is unique. Consider the language
  whose executions write unique values, with the following predicates:
  \begin{align*}
    & \mathsf{w}(e,o) & \text{$o$ is a {\sf write} operation in $e$, and}, \\
    & \mathsf{v}(e,o_1,o_2) & \text{$o_1$ and $o_2$ read/write the same values.}
  \end{align*}
  We define the function $G(x_\mathsf{w}, y_\mathsf{w}, z_\mathsf{v})$ over
  valuations to the predicates above to be satisfied if and only if:
  \begin{align*}
    x_\mathsf{w} \land z_\mathsf{v}
  \end{align*}
  Intuitively, this defines a simple matching scheme for which write
  operations match themselves, and read operations match the write which wrote
  the value read. In the case such a write exists, it is unique in any
  execution which writes unique values. The match is otherwise undefined. This
  scheme is exactly that which is specified in Example~\ref{ex:matching} of
  Section~\ref{sec:histories}.

@@

In what follows, we denote the operations of an execution $e$ by $O_e$.
To consider whether a given pair $o_1, o_2$ of operations of an execution $e$
is a match according to the simple matching scheme based on ${\sf g}$, for
the given language $L$, we define the macro $\textsc{IsMatch}_{e,o_1,o_2}$:
\begin{align*}
  \mathsf{g}\big( P(e,o_1), P(e,o_2), Q(e,o_1,o_2) \big).
\end{align*}
To enforce that each operation of an execution $e$ will have a unique match,
we define the macro $\textsc{UniqueMatch}_e$:
\begin{align*}
  \bigwedge_{o_1,o_2 \in O_e} \Big(
    \textsc{IsMatch}_{e,o_2,o_2} \implies
    \bigwedge_{o_1' \neq o_1} \lnot \textsc{IsMatch}_{e,o_1',o_2}
  \Big).
\end{align*}
Then a given set $O \subseteq O_e$ constitutes a match according to
${\sf g}$ when all operations $o_2 \in O$ target some operation $o_1 \in O$,
and no other operation $o_2 \in O \setminus O_e$ does. We express this with the
macro $\textsc{EntireMatch}_{e,O}$:
\begin{align*}
  \bigvee_{o_1 \in O}
  \Big(
  \bigwedge_{o_2 \in O} \textsc{IsMatch}_{e,o_1,o_2}
  \land \bigwedge_{o_2 \in O_e \setminus O} \lnot \textsc{IsMatch}_{e,o_1,o_2}
  \Big)
\end{align*}
Finally, given a pair $\tup{e,e'} \in F$, we prohibit the operations $O_e
\setminus O_{e'}$ from constituting a match according to ${\sf g}$, since if
${\sf g}$ normalized $\mathcal{I}$, and $e \in \mathcal{I}$, then $e'$ should
also be in $\mathcal{I}$. We express this exclusion for all pairs of $F$
with the macro $\textsc{Normalizes}_{F}$:
\begin{align*}
  \bigwedge_{\tup{e,e'}\in F} \Big(
    \textsc{UniqueMatch}_{e}
    \land \lnot \textsc{EntireMatch}_{e,(O_e \setminus O_{e'})}
  \Big).
\end{align*}
We thus check satisfiability for the conjunction of non-matches $O_e \setminus
O_{e'}$ over all pairs $\tup{e,e'} \in F$.

@@

  Consider again the language $L_\mathrm{reg}$ of read and write methods of
  Example~\ref{ex:register-language}, yet this time without the predicate ${\sf
  v}(e,o_1,o_2)$. The following table lists all possible predicate valuations
  $\tup{x_{\sf w}, y_{\sf w}}$, and for each valuation a valid execution which
  excludes the positive valuation of ${\sf g}(x_{\sf w}, y_{\sf w})$ in the
  satisfaction of $\textsc{Normalize}_F$.
  \begin{center}
    \begin{tabular}{cc|ll}
      $x_{\sf w}$ & $y_{\sf w}$ & counterexample & reason \\
      \hline
      $0$ & $0$ & ${\rm w}(1)\ {\rm r}(1)\ {\rm r}(1)$ & not unique \\
      $0$ & $1$ & $\underline{{\rm w}(1)\ {\rm r}(1)}\ {\rm r}(1)$ & $\to_\mathrm{m}$ \\
      $1$ & $0$ & ${\rm w}(1)\ {\rm w}(2)\ {\rm r}(2)$ & not unique \\
      $1$ & $1$ & ${\rm w}(1)\ {\rm w}(2)$ & not unique \\
    \end{tabular}
  \end{center}
  Thus any enumeration $F$ including the executions above excludes every
  possible valuation of $\tup{x_{\sf w}, y_{\sf w}}$, resulting in an
  unsatisfiable $\textsc{Normalize}_F$.

>>>>

\section{Matching Scheme Inference}
\label{sec:matching}

@@

The required manual specification includes identifying a class of executions to
which an implementation should be exposed, and a set of predicates
required in the logic of matching schemes. Formally, a \emph{language} $L =
\tup{E,P,Q}$ is a set $E$ of executions, along with finite sequences $P = P_1
P_2 \ldots$ and $Q = Q_1 Q_2 \ldots$ of binary and ternary predicates
$P_i(e,o_1)$ and $Q_i(e,o_1,o_2)$ ranging over executions and their
operations. When $P$ is a $k$-length sequence of $n$-ary predicates, we write
$P(x_1, \ldots, x_n)$ to denote the valuation sequence
\begin{align*}
  P_1(x_1, \ldots, x_n) \ldots P_k(x_1, \ldots, x_n).
\end{align*}
Given a language $L = \tup{E,P,Q}$, we say a matching scheme $M$ is
\emph{simple} when there exists an $n$-ary Boolean function $G$, for $n =
2\cdot|P|+|Q|$, such that for each execution $e \in E$
\begin{itemize}

@@

  We say an execution of read and write methods \emph{writes unique values} if
  the argument value to each write operation is unique. Consider the language
  whose executions write unique values, with the following predicates:
  \begin{align*}
    & \mathsf{w}(e,o) & \text{$o$ is a {\sf write} operation in $e$, and}, \\
    & \mathsf{v}(e,o_1,o_2) & \text{$o_1$ and $o_2$ read/write the same values.}
  \end{align*}
  We define the function $G(x_\mathsf{w}, y_\mathsf{w}, z_\mathsf{v})$ over
  valuations to the predicates above to be satisfied if and only if
  $x_\mathsf{w} \land z_\mathsf{v}$.
  Intuitively, this defines a simple matching scheme for which write
  operations match themselves, and read operations match the write which wrote
  the value read. In the case such a write exists, it is unique in any
  execution which writes unique values. The match is otherwise undefined. This
  is the scheme specified in Example~\ref{ex:matching} of
  Section~\ref{sec:histories}.

@@

In what follows, we denote the operations of an execution $e$ by $O_e$.
To consider whether a given pair $o_1, o_2$ of operations of an execution $e$
is a match according to the simple matching scheme based on ${\sf g}$, for
the given language $L$, we define the macro $\textsc{IsMatch}_{e,o_1,o_2}$:
\begin{align*}
  \mathsf{g}\big( P(e,o_1), P(e,o_2), Q(e,o_1,o_2) \big).
\end{align*}
To enforce that each operation of an execution $e$ will have a unique match,
we define the macro $\textsc{UniqueMatch}_e$:
\begin{align*}
  \bigwedge_{o_1,o_2 \in O_e} \Big(
    \textsc{IsMatch}_{e,o_1,o_2} \implies
    \bigwedge_{o_1' \neq o_1} \lnot \textsc{IsMatch}_{e,o_1',o_2}
  \Big).
\end{align*}
Then a given set $O \subseteq O_e$ constitutes a match according to
${\sf g}$ when all operations $o_2 \in O$ target some operation $o_1 \in O$,
and no other operation $o_2 \in O \setminus O_e$ does. We express this with the
macro $\textsc{EntireMatch}_{e,O}$:
\begin{align*}
  \bigvee_{o_1 \in O}
  \Big(
  \bigwedge_{o_2 \in O} \textsc{IsMatch}_{e,o_1,o_2}
  \land \bigwedge_{o_2 \in O_e \setminus O} \lnot \textsc{IsMatch}_{e,o_1,o_2}
  \Big)
\end{align*}
Finally, given a pair $\tup{e,e'} \in F$, we prohibit the operations $O_e
\setminus O_{e'}$ from constituting a match according to ${\sf g}$, since if
${\sf g}$ normalized $\mathcal{I}$, and $e \in \mathcal{I}$, then $e'$ should
also be in $\mathcal{I}$. We express this exclusion for all pairs of $F$
with the macro $\textsc{Normalizes}_{F}$:
\begin{align*}
  \bigwedge_{\tup{e,e'}\in F} \Big(
    \textsc{UniqueMatch}_{e}
    \land \lnot \textsc{EntireMatch}_{e,(O_e \setminus O_{e'})}
  \Big).
\end{align*}
We thus check satisfiability for the conjunction of non-matches $O_e \setminus
O_{e'}$ over all pairs $\tup{e,e'} \in F$.

@@

  Consider again the language $L_\mathrm{reg}$ of read and write methods of
  Example~\ref{ex:register-language}, yet this time without the predicate ${\sf
  v}(e,o_1,o_2)$. The following table lists valid executions excluding each
  of the possible predicate valuations $\tup{x_{\sf w}, y_{\sf w}}$
  in the satisfaction of $\textsc{Normalize}_F$.
  \begin{center}
    \begin{tabular}{cc|ll}
      $x_{\sf w}$ & $y_{\sf w}$ & counterexample & reason \\
      \hline
      $0$ & $0$ & ${\rm w}(1)\ {\rm r}(1)\ {\rm r}(1)$ & not unique \\
      $0$ & $1$ & $\underline{{\rm w}(1)\ {\rm r}(1)}\ {\rm r}(1)$ & $\to_\mathrm{m}$ \\
      $1$ & $0$ & ${\rm w}(1)\ {\rm w}(2)\ {\rm r}(2)$ & not unique \\
      $1$ & $1$ & ${\rm w}(1)\ {\rm w}(2)$ & not unique \\
    \end{tabular}
  \end{center}
  Thus any enumeration $F$ including the executions above results in an
  unsatisfiable $\textsc{Normalize}_F$.
